### 2016-05-20 Paper_v1.tex
May19

\subsection*{Sensitivity analysis}
We conducted a local sensitivity analysis to determine which of the kinetic and control parameters affected model performance.
We calculated performance as area under the CAT curve, which was directly related to CAT synthesis rate, as the culture time was fixed and no CAT degradation was modeled.
We randomly chose 180 sets of the 18,000 sets in the ensemble and defined these as our sub-ensemble; for each set in the sub-ensemble, we varied the rate constant and saturation constants of each metabolic flux and measured the resulting change in CAT production to estimate the sensitivity of model performance to that parameter.
We did the same for the control parameters, both the order (Hill coefficient) and gain (related to the the dissociation constant).
This allowed us to estimate the relative importance of the kinetic and control parameters to CAT production across the ensemble.
Of the rate constants, those with the highest positive sensitivities were CAT synthesis, GTP synthesis, GMP synthesis, glutamine synthesis, and aspartate synthesis (Fig.~\ref{fig:LocalSensitivity}, top).
This is explained by GTP and the amino acids being reactants for CAT synthesis.
The rate constants with the largest negative sensitivities were GTP degradation, arginine synthesis, and UMP synthesis.
While GTP degradation is obvious, the others can be explained in that they consume ATP as well as several amino acids, all of which are reactants for CAT synthesis.
Of the saturation constants, the reverse is seen: the largest positive sensitivities are those associated with arginine synthesis, while the largest negative sensitivities are those associated with CAT synthesis, GTP synthesis, and GMP synthesis (Fig.~\ref{fig:LocalSensitivity}, middle).
This is because an increase in saturation constant causes a decrease in the corresponding reaction rate \cite{?}.
The control parameters were seen to be the least significant and the most uncertain (Fig.~\ref{fig:LocalSensitivity}, bottom).
Only two had a small standard error across the ensemble, relative to the ensemble mean sensitivity: the gain and order for pyruvate acting as an inhibitor on the pdh reaction.
This could be because pdh consumes pyruvate and diverts carbon away from the pathways that ultimately contribute to CAT production.
Taken together with the lack of change in glucose uptake and CAT production when control is removed (Fig.~\ref{fig:BothCarbon}), this suggests that allosteric control is relatively unimportant to CAT production and metabolism as a whole compared to kinetic effects.

@@

\subsection*{Calculation of CAT yield}
We calculated the carbon yield of CAT production for our experimental data and our best-fit parameter set as a function of the initial and final concentrations and the carbon numbers of CAT, glucose, and amino acids.
Arginine and glutamate were excluded due to not being present in the Swartz dataset.
The experimental data displayed a CAT yield of 0.0865, while the best-fit parameter set displayed a CAT yield of 0.0871.
We then used sequence-specific FBA to calculate a theoretical maximum CAT yield of 0.1942.
Thus, we showed that our experimental dataset and best-fit parameter set were producing CAT at 44.5\% and 44.9\% of the theoretical maximum, respectively.
This allowed us to quantify the amount of carbon being diverted to byproducts, and suggests that there is potential for a doubling of CAT production by reducing this diversion of carbon.

@@

The discussion has three (sometimes four) paragraphs:
\begin{enumerate}
	\item{\textbf{First~paragraph}: Present a modified version of the last paragraph of the introduction. In this study, [...]. Taken together, [killer statement]}
	\item{\textbf{Second~paragraph}: Contrast the key findings of the study with other computational/experimental studies}
	\item{\textbf{Third~paragraph}: Present future directions. If you had more time, what would like to do? Highlight the key shortcomings of the approach and how will we address them in the future.
	In this case, we will have a scaling issue if we extend to genome scale. We should extend to dynamic cases, and we need to experimentally validate the findings.}
\end{enumerate}

@@

% Figures and captions go here ...
\begin{figure}[ht]
\centering
\includegraphics[width=0.75\textwidth]{./Figures/Network.pdf}
\caption{Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and NADPH/NADH transfer. Mean ± standard error for the reaction flux at 1.5 hrs (top) and 3 hrs (bottom), normalized to CAT synthesis flux.}
\label{fig:Network}
\end{figure}

>>>>

\subsection*{Sensitivity analysis}
We conducted a local sensitivity analysis to determine which of the kinetic and control parameters affected model performance.
We calculated performance as area under the CAT curve, which was directly related to CAT synthesis rate, as the culture time was fixed and no CAT degradation was modeled.
We randomly chose 180 sets of the 18,000 sets in the ensemble and defined these as our sub-ensemble; for each set in the sub-ensemble, we varied the rate constant and saturation constants of each metabolic flux and measured the resulting change in CAT production to estimate the sensitivity of model performance to that parameter.
We did the same for the control parameters, both the order (Hill coefficient) and gain (related to the the dissociation constant).
This allowed us to estimate the relative importance of the kinetic and control parameters to CAT production across the ensemble.
Of the rate constants, those with the highest positive sensitivities were CAT synthesis, GTP synthesis, GMP synthesis, glutamine synthesis, and aspartate synthesis (Fig.~\ref{fig:LocalSensitivity}, top).
This is explained by GTP and the amino acids being reactants for CAT synthesis.
Also, GMP synthesis increases the total amount of guanosine, allowing for more GTP production.
The rate constants with the largest negative sensitivities were GTP degradation, arginine synthesis, and UMP synthesis.
While GTP degradation is obvious, the others can be explained in that they consume ATP as well as several amino acids, all of which are reactants for CAT synthesis.
Of the saturation constants, the reverse is seen: the largest positive sensitivities are those associated with arginine synthesis, while the largest negative sensitivities are those associated with CAT synthesis, GTP synthesis, and GMP synthesis (Fig.~\ref{fig:LocalSensitivity}, middle).
This is because an increase in saturation constant causes a decrease in the corresponding reaction rate \cite{?}.
The control parameters were seen to be the least significant and the most uncertain (Fig.~\ref{fig:LocalSensitivity}, bottom).
Only two had a small standard error across the ensemble, relative to the ensemble mean sensitivity: the gain and order for pyruvate acting as an inhibitor on the pdh reaction.
This could be because pdh consumes pyruvate and diverts carbon away from the pathways that ultimately contribute to CAT production.
Taken together with the lack of change in glucose uptake and CAT production when control is removed (Fig.~\ref{fig:BothCarbon}), this suggests that allosteric control is not the limiting factor to CAT production.

@@

\subsection*{Calculation of CAT yield}
We calculated the carbon yield of CAT production for our experimental data and our best-fit parameter set as a function of the initial and final concentrations and the carbon numbers of CAT, glucose, and amino acids.
Arginine and glutamate were excluded due to not being present in the Swartz dataset.
The experimental data displayed a CAT yield of 0.0865, while the best-fit parameter set displayed a CAT yield of 0.0871.
We then used sequence-specific FBA to calculate a theoretical maximum CAT yield of 0.1942.
Thus, we showed that our experimental dataset and best-fit parameter set were each producing CAT at 45\% of the theoretical maximum.
This allowed us to quantify the amount of carbon being diverted to byproducts, and suggests that there is potential for a doubling of CAT production by reducing this diversion of carbon.

@@

%The discussion has three (sometimes four) paragraphs:
%\begin{enumerate}
%	\item{\textbf{First~paragraph}: Present a modified version of the last paragraph of the introduction. In this study, [...]. Taken together, [killer statement]}
%	\item{\textbf{Second~paragraph}: Contrast the key findings of the study with other computational/experimental studies}
%	\item{\textbf{Third~paragraph}: Present future directions. If you had more time, what would like to do? Highlight the key shortcomings of the approach and how will we address them in the future.
%	In this case, we will have a scaling issue if we extend to genome scale. We should extend to dynamic cases, and we need to experimentally validate the findings.}
%\end{enumerate}

We constructed a model of cell-free \textit{E. coli} metabolism with CAT production based on saturation kinetics and allosteric control.
We trained the model against measurements of glucose, CAT, organic acid intermediates, energy species, and amino acids.
We generated an ensemble of 18,000 parameter sets that predict the experimental dataset, with the exception of some amino acids.
We calculated the carbon yield of CAT production at 45\% of the theoretical maximum obtained by flux balance analysis; this was in line with the CAT yield shown by the dataset.
We conducted sensitivity analyses and determined that CAT production was most sensitive to the rate constants, saturation constants, and enzyme initial conditions associated with CAT synthesis, GTP synthesis, GMP synthesis, and amino acid synthesis, and to a lesser extent some amino acid initial conditions.
CAT production was much less sensitive to control parameters, suggesting that while allosteric control affects some organic acid intermediates, it does not limit CAT production.
The initial conditions of glucose and oxygen were even less important to CAT production, suggesting that it can be driven by other means.
However, the low sensitivities to glucose and oxygen only apply in the specific case of the parameter sets that were studied.
When other species' initial conditions and associated parameters are different, glucose and oxygen can become more important to CAT production.

2nd paragraph: Compare to other studies
This is the first study to model a cell-free \textit{E. coli} system at this scope, toward genome-scale.

One important direction for the future is to expand the biological description, especially with regard to amino acids.
Specifically, adding more reactions that consume amino acids would improve the model's ability to predict those that show a decrase in the experimental data.
Also, including specific transcription and translation steps for CAT would allow us to more accurately model the complexity and the resource cost of protein synthesis.
Another area for future work is to more thoroughly sample parameter space.
For the metabolites in the dataset, initial conditions were fixed at the initial data values.
All other parameters were varied in a manner so as to best fit the dataset.
However, the resulting ensemble may not represent every biological or practical possibility.
In a different region of parameter space, the system could behave differently, including the flux distribution through the network, the accuracy and spread of ensemble fits, the relative sensitivities, and the yield as a percentage of the theoretical maximum.
Testing the model under a variety of conditions could strengthen or challenge the findings of this study.
Further experimentatiion could also be used to gain a deeper understanding of model performance under a variety of conditions.
Specifically, CAT production performed in the absence of amino acids could inform the system's ability to manufacture them, while experimentation in the absence of glucose or oxygen could shed light on how important they are to protein synthesis, and under which conditions.
Finally, the approach should be extended to other protein products.
CAT is only a test protein used for model identification; the modeling framework, and to some extent the parameter values, should be protein agnostic.
An important extension of this study would be to apply its insights to other protein applications, where possible.

@@

% Figures and captions go here ...
\begin{figure}[ht]
\centering
\includegraphics[width=0.75\textwidth]{./Figures/Network.pdf}
\caption{Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and NADPH/NADH transfer. FBA flux value (top), and mean ± standard error across ensemble at 1.5 hrs (middle) and 3 hrs (bottom), normalized to CAT synthesis flux.}
\label{fig:Network}
\end{figure}
